---
title: "Color schemes in accordance with the ICAE corporate design"
author: "Claudius Gräbner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This package provides some functions that make it easy to create figures in accordance with the corporate design colors of the Institute for Comprehensive Analysis of the Economy (ICAE) at the Johannes Kepler University in Linz.^[More information about the ICAE can be found [here](https://www.jku.at/en/institute-for-comprehensive-analysis-of-the-economy/).]

The package is still under development. Here is just a quick overview over its functions. Note that all exported function already have a complete documentation so you might refer to the help function for more details.

For the following porpuses, this example code and the packages `dplyr` as well as `ggplot2` are used:
```{r, echo=TRUE, message=FALSE}
x <- seq(1,10, length.out=30)
y <- rep(1, length(x))
data <- data.frame(x=x, y=y)
palettes <- c("main", "cool", "hot", "mixed")
library(dplyr)
library(ggplot2)
library(icaeDesign)
```

# The palettes

Currently, the package contains the following four palettes:

```{r, fig.height=0.5, echo=FALSE}
pal_plots <- list()
for (n in palettes){
  pal_name <- n
  pal_plots[[pal_name]] <- ggplot(data, aes(x, y, z= x)) + geom_tile(aes(fill = x)) + 
  scale_fill_icae(pal_name, discrete = F) +
  ggtitle(paste0(pal_name)) +
  theme_void() + theme(legend.position = "none", 
                       plot.title = element_text(hjust = 0.5, size = 10))
  print(pal_plots[[pal_name]] )
}
```
  

All of them come with both continuous and discrete versions.

## Using the palettes

The key functions are `scale_color_icae()` and `scale_fill_icae()`, which allow to control the `color` and `fill` aesthetics in ggplot2 objects respectively. Please note that currently only the application to ggplot2 objects is supported.

In case of a barplot you might use `scale_fill_icae()` as follows:

```{r, message=FALSE, fig.width=4}
data("iris")
iris %>%
  group_by(Species) %>%
  summarise_all(mean) %>%
  ungroup() %>%
  ggplot(.) + geom_bar(aes(x=Species, y=Petal.Width, fill=Species), stat = "identity") + 
  scale_fill_icae() + theme_minimal()
```

Note that by default the functions return a discrete scale and for continuous cases you need to set `discrete = FALSE` explicitly.

```{r, fig.width=4}
iris %>%
  ggplot(.) + geom_point(aes(x=Sepal.Length, color=Sepal.Width, y=Petal.Length)) +
  scale_color_icae(discrete = F) + theme_minimal()
```

For distinguishing discrete categories via color, such as countries, I found the `mixed` palette superior to `main`:

```{r, fig.width=4}
iris %>%
  ggplot(.) + geom_point(aes(x=Sepal.Width, y=Sepal.Length, color=Species)) +
  scale_color_icae(palette = "mixed", discrete = T) + theme_minimal()
```

# The new ggplot theme

The package also features a new theme for ``ggplot2`` objects, called ``theme_icae``.
It includes a number of changes that I found useful in improving plot appearance in general, but it is not specifically built to fit the ICAE color scheme.

```{r, fig.width=4}
classic_theme_plot <- iris %>%
  ggplot(.) + geom_point(aes(x=Sepal.Width, y=Sepal.Length, color=Species)) +
  ggtitle("The standard ggplot theme") +
  scale_color_icae(palette = "mixed", discrete = T)

icae_theme_plot <- iris %>%
  ggplot(.) + geom_point(aes(x=Sepal.Width, y=Sepal.Length, color=Species)) +
  ggtitle("The theme_icae() appearance") +
  scale_color_icae(palette = "mixed", discrete = T) + theme_icae()

ggpubr::ggarrange(classic_theme_plot, icae_theme_plot, ncol = 2, nrow = 1, common.legend = F)
```
